A Possibilistic Valid-time Model

### 2011-12-12 trunk/ipmu2012/introduction.tex
Last temporary changes... this evening final changes.

First of all, the concept of \emph{time} itself is very complex. Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}. More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{devos98,Cru97}. Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.

Next to the inherent difficulties of time, an issue is \emph{temporal granularity}. A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application. For example, the age of an adult human being is usually expressed in years: one will use sentences like `Laura is 21 years old' instead of `Laura is 21 years, 3 months and 4 days old'. In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.

Another issue is the possible inherent imperfection in the description of notions of time. The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression. The temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D. and 1400 A.D.' contains imperfection caused by the inaccuracy in the given temporal notion, though precise boundaries are given.

The transition between granularities~\cite{Lin97} is also considered as a source of imprecision. In a sentence like `The Alhambra is 612 years old', the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity. However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed. Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.

To allow information systems to cope with imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, though some other tools, like rough sets, have also been used~\cite{Qia09}. The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.

Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more problems, as the modelling of a time-variant object has an impact on the database consistency. Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95}, but all of these consider time as a crisp notion. In recent works though, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,fuzz2009}.

Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval. There are several proposals for this transformation, but some, like those presented in~\cite{fuzz2009} are misleading.

In this work we will propose a possibilistic valid time model for relational databases. The base of the model is the possibilistic framework presented in~\cite{Pon11}. This framework deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.

The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and temporal databases are presented. 
Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.
Finally, section \ref{sec:futher-research} presents the main conclusions and some lines for future work in this research.

>>>>

First of all, the concept of \emph{time} itself is very complex. More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}. Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.

%Next to the inherent difficulties of time, an issue is \emph{temporal granularity}. A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application. For example, the age of an adult human being is usually expressed in years: one will use sentences like `Laura is 21 years old' instead of `Laura is 21 years, 3 months and 4 days old'. In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.

One issue in these attempts at temporal modelling is off course the inherent complexities of the concept of time. Another is the possible inherent imperfection in the description of notions of time. The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression. The temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D. and 1400 A.D.' contains imperfection caused by the inaccuracy in the given temporal notion, though precise boundaries are given. The transition between temporal granularities~\cite{Lin97} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{Cru97}.

%The transition between granularities~\cite{Lin97} is also considered as a source of imprecision. In a sentence like `The Alhambra is 612 years old', the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity. However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed. Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.

To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,schockaert08}.

Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency. Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion. In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.

Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval. There are several proposals for this transformation, but some, like those presented in~\cite{Garrido2009} are misleading.

In this work we will propose a possibilistic valid time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.

The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.
Finally, section \ref{sec:futher-research} presents the main conclusions and some lines for future work in this research.

### 2011-12-12 trunk/ipmu2012/proposal.tex
Last temporary changes... this evening final changes.

The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.

\subsection{Representation of ill-known valid-time intervals}
Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.

\begin{definition}
A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
\begin{equation}
PVP = \left[X,\ Y \right] 
\end{equation}
$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
\end{definition}

The possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}. This representation allows overlapping (Fig. \ref{fig:pvp}).

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVT}.}
  \label{fig:pvp}
\end{figure}

\subsection{Storage of valid-time intervals}
Each database row containing a \emph{PVP} stores it as two triangular possibility distributions. In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98}. In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):

\begin{itemize}
\item
\emph{NULL}: This constant refers to a completely ignorance about the value. The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
\item
\emph{UNKNOWN}: The point has a value but it is unknown. The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
\item
\emph{UNDEFINED}: The point does not have a value. The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
\end{itemize}

\begin{table}
\caption{Relational representation for a Possibilistic Valid-Time Period.}
\centering
\begin{tabular}{c c c c c c}
\hline
Value & FT & F1 & F2 & F3  \\ \hline
UNKNOWN & 0 & NULL & NULL & NULL  \\ 
UNDEFINED & 1 & NULL & NULL & NULL  \\ 
NULL & 2 & NULL & NULL & NULL  \\ 
$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ \\ 
\hline
\end{tabular}
\label{table:relational-representation-pvp}
\end{table}

\subsection{Querying ill-known valid-time intervals}

\subsubsection{Query specification}

\subsubsection{Query evaluation}

\subsubsection{Ranking}

>>>>

In this section, a Valid-time model will be presented. The focus will off course be on the handling of imperfect Valid-time notions. First, the representation and storage of Valid-time will be presented. Next, a way of querying the model is proposed. Finally, an example is given.
%The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.

\subsection{Valid-time Intervals}
\subsubsection{Representation}
\label{subsec:representation-ill-known}
A database models a real object by storing the object's value for each attribute describing a property of the object. Thus, a Valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic Valid-time Period} to the record describing said object state:

\begin{definition}
A \emph{Possibilistic Valid-time Period} (PVP) is an ill-known interval in time, which models a time period during which an object is valid.
\end{definition}

Off course, the object of which the PVP models the Valid-time is the object in the particular state modelled by the record that is associated with the PVP. Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start or end point of a time interval (and thus about the time interval itself) if such uncertainty exists. The interpretation is \emph{disjunctive}: the PVP represents exactly one Valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown. In the presented model, only PVPs are considered in which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have the exact same characteristics of the membership functions of fuzzy numbers. A perfectly known start or end point can then be modelled by such an ill-known value defined by a possibilistic variable $P$ for which $\exists ! x : \mu_{P}(x) > 0$.

As mentioned in \cite{Pon11}, this approach differs from the one where a Valid-time period is represented by one fuzzy set. This one fuzzy set is seen as a possibility distribution on $\mathbb{R}$. However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\wp(\mathbb{R})$.

%The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.

%\begin{definition}
%A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
%\begin{equation}
%PVP = \left[X,\ Y \right] 
%\end{equation}
%$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
%\end{definition}

%For convenience, the possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}. This representation allows overlapping (Fig. \ref{fig:pvp}). Note that while the two ill-known values $X,Y \in \mathbb{R}$, the fuzzy interval  $[X,Y] \in \mathbb{R}^2$.

%\begin{figure}[h!]
%  \centering
%  \input{./graphs/2_triangular.tex}
%  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVP}.}
%  \label{fig:pvp}
%\end{figure}

\subsubsection{Storage}
\label{subsec:storage}
To store a PVP, the possibility distributions defining the ill-known start and end point are stored. To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98} is used. Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants. To store an ill-known point, four values (FT, F1, F2 and F3) are stored. They are explained in table \ref{table:relational-representation-pvp}. Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database.

%\begin{itemize}
%\item
%\emph{NULL}: This constant refers to a completely ignorance about the value. The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
%\item
%\emph{UNKNOWN}: % The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
%\item
%\emph{UNDEFINED}: The point does not have a value. %The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
%\end{itemize}
%

\begin{table}
\centering
\begin{tabular}{c c c c c c l p{2cm}}
\hline
Value & FT & F1 & F2 & F3 & $\mu(x)$ & Description \\ \hline
UNKNOWN & 0 & N & N & N  & $1$ & Any value is equally possible\\ 
UNDEFINED & 1 & N & N & N & $0$ & The value is not defined in the \\
          &   &   &   &   &     & attribute domain\\ 
NULL & 2 & N & N & N &not defined & Nothing is known about the value \\ 
$\left[D,\ a,\ b \right]$ & 3 & $D$ & $a$ & $b$ & $\mu(\left[D,\ a,\ b \right])$ & Ill-known value \\ 
\hline 
\end{tabular}
\label{table:relational-representation-pvp}

\vspace{10pt}

\caption{Relational representation for an ill-known time point. FT denotes Fuzzy Type. Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $\left[D,\ a,\ b \right]$. Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
\end{table}

\vspace{-25pt}

\subsection{Querying Ill-known Valid-time Intervals}
In order to present a complete model, the tools for querying will be provided in this subsection. The focus will evidently lie on the querying of Valid-time periods. First, the query structure will be defined, followed by the evaluation and ranking methods.

%In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.

%It is important to notice that every Valid-time notion in the database is represented using a PVP. Thus, Valid-time intervals in the database can be partially unknown.

%In the query, the user is allowed to specify both non-temporal preferences and a temporal constraint.

%This allows the user to specify both the preferences and an ill-known valid-time interval in the query. It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known). In the query specification, the user is allowed to express a crisp time interval.

\subsubsection{Query Structure}
It is important to notice that every Valid-time notion in the database is represented using a PVP. Thus, Valid-time intervals in the database can be partially unknown. In this framework, a query has two separate parts:

% one is a temporal constraint and the other is the collection of query specifications for regular attributes:

%the first one is the temporal specification. The second part is the query specification for regular attributes.

\begin{definition}
A query $\tilde Q$ is specified by:
\begin{equation}
\label{eq:query-definition}
\tilde Q = \left( Q^{time}, Q \right)
\end{equation}
\end{definition}
Here, $Q$ denotes the collection of (possibly fuzzy) non-temporal preferences of the user and $Q^{time}$ denotes the temporal constraint given by the user:
\begin{definition}
 $Q^{time}$ is defined by:
\begin{equation}
Q^{time} = \left( I , AR \right)
\end{equation}
\end{definition}
Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations\cite{Allen83}. The interpretation is that for a record with Valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.

%In table \ref{tab:allen-relations}, .

%\vspace{-35pt}

\subsubsection{Query evaluation}
In fuzzy querying of regular (relational) databases, the modelling of query satisfaction is a matter of degree. Usually, the evaluation of the query requirements for a record results in a satisfaction degree $s$, where $s$ lies in $\left[0,1\right]$, where 0 denotes total dissatisfaction and 1 denotes complete satisfaction. In crisp querying, the evaluation of query requirements for a record results in the accepting or rejecting of the record as a part of the result set. This can be modelled using satisfaction degrees, by assigning rejection a degree of $0$ and acceptance a degree of $1$ and not using any other value in $\left[0,1\right]$.

The evaluation of a query $\tilde{Q} = \left( Q^{time}, Q \right)$, where $Q^{time} = \left( I, AR \right)$ is now handled as follows. For each record $r$ in the database, with the Valid-time notion of $r$ specified by PVP $J$, two things happen independently:

% by a combination of non-temporal query requirements $Q$ is noted $e_{Q}(r) \in \left[0,1\right]$ in this work. The total query evaluation method is now the following.  and Allen relation $AR$
\begin{itemize}
\item
The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$. The model presented here accepts any way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$. 
\item
Depending on AR, a specific set of ill-known constraints is considered. For each AR, this set can be found in table \ref{tab:allen-relations}. The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \ref{ill-known-pos} and \ref{ill-known-nec}. The sum of these possibility and necessity scores gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$. Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural score.
\end{itemize}

\begin{table}[h]
\label{tab:allen-relations}
\centering
\begin{tabular}{|c|c|c|}
\hline
Allen Relation & Constraints & $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
\hline
I before J & $C_1 \triangleq \left(<,X\right)$ & $\sup_{a>w}\pi_x(w)$\\
\hline

\multirow{2}{*}
{I equal J} & $C_1\triangleq \left(\geq,X\right)$,$C_2\triangleq \left(\neq,X\right)$ & $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$ \\
 & $C_3\triangleq \left(\leq,Y\right)$,$C_4\triangleq \left(\neq,Y\right)$ & $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
\hline

I meets J & $C_1\triangleq \left(\leq,X\right)$ ,$C_2\triangleq \left(\neq,X\right)$ & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
\hline

\multirow{2}{*}
{I overlaps J} & $C_1\triangleq \left(<,Y\right)$,$C_2\triangleq \left(\leq,X\right)$ & $\min ( \sup_{b>w}\pi_Y(w), $ \\
 & $C_3\triangleq \left(\geq,X\right)$ & $\sup_{a \geq w}\pi_X(w),\sup_{a \leq w}\pi_X(w))$ \\
\hline

\multirow{2}{*}
{I during J} & $C_1\triangleq \left(>,X\right)$, $C_2\triangleq \left(\leq,Y\right)$ & $\max ( \min ( \sup_{a<w}\pi_X(w),\sup_{b \geq w}\pi_Y(w)),$ \\
 & $C_3\triangleq \left(\geq,X\right)$ ,$C_4\triangleq \left(<,Y\right)$ & $\min ( \sup_{a \leq w }\pi_X(w),\sup_{b>w}\pi_Y(w)$\\
\hline

{I starts J} & $C_1\triangleq \left(\geq,X\right)$,$C_2\triangleq \left(\neq,X\right)$  & $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
\hline

{I finishes J} & $C_1\triangleq \left(\leq,Y\right)$, $C_2\triangleq \left(\neq,Y\right)$  & $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$ \\
\hline

\end{tabular}

\vspace{10pt}

\caption{Allen's relations represented in the framework. Here, $I = \left[a, b\right]$ denotes the PVP in the query, $J = \left[X, Y\right]$ denotes the PVP of the record, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively. For each of Allen's relations $AR$, the corresponding value in column `Constraints' gives the constraints corresponding to $AR$. The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints' and thus that $I$ $AR$ $J$ holds. These formulas are based on equation \ref{ill-known-pos}.}

\vspace{-25pt}

\end{table}

\subsubsection{Ranking and aggregation}
In order to present the results to the user, a ranking method is used. First of all, $e_{Q^{time}}(r)$ is rescaled to the unit interval. The final ranking $e_{final}$ is now given by a convex combination:

%the possibility and the necessity are ranked with the convex combination in \eqref{eq:convex-comb} with $\omega=0.5$. Finally this value is aggregated with the rest of the criteria with the same combination. %In this case the value for the aggregation is also $\omega = 0.5$

\begin{equation}
\label{eq:convex-comb}
e_{final}(r)\ =\ \omega*e_{Q}(r)\ +\ (1-\omega)*e_{Q^{time}}(r), \omega \in \left[0, 1 \right]
\end{equation}

The use of this convex combination allows a record to make up for a low score on the temporal constraint by a good score on the non-temporal constraint (or vice versa). Changing $\omega$ also allows granting the temporal constraint more weight with respect to the non-temporal constraint (or vice versa).

%By increasing the value for the parameter $\omega$, the preferences expressed in the query $Q$ can be given more importance. By lowering the value for $\omega$, the temporal criteria is emphasized.
%The following example illustrates the querying process.

\subsection{Example}

FROM HERE

%\begin{example} 
\subsubsection{The database}
Consider a database describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute containing the approximate period of time during which the car model was sold. The temporal data is stored as explained in subsection \ref{subsec:storage}. The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer, for the convenience of the representation. Therefore the chronons in our example will be years. The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sample table for the car models. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht]
\centering
\begin{tabular}{c c c c c c c}
\hline
ID & IID & Segment & Manufacturer & Name & Start & End  \\ [0.5ex]
\hline
001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
003 & 1 & B & Citroen & C2 & [2002,2,2] & [2006,1,1] \\
001 & 2 & B & Peugeot & 206 & [2000,1,2] & [2011,2,1] \\
001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
\hline
\end{tabular}
\vspace{10pt}
\caption{Sample database}
\label{tb:car-models}
\end{table}

\textbf{The query}
Consider the following query:

\emph{The user wants to obtain a list of models in the segment B for the manufacturer Peugeot before the year interval 2001-2005.}

The query is translated to the following expression, using equation \eqref{eq:query-definition}. The evaluation for the criteria are presented in the result table \ref{tb:results}.

\begin{center}
$\tilde{Q} = \left(  c^{time}, c_{segment} \right)$
\end{center}

Where:
\begin{itemize}
\item
$c^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
\item
$c_{segment}\ = \ $ Peugeot.
\end{itemize}

\begin{table}[ht]
\caption{Result table and ranking}
\centering
\begin{tabular}{c c c c c c c}
\hline
ID & IID &  $\Pos_{c^{time}}$ &$\Nec_{c^{time}}$ & $e(C^{time})$ & $c_{segment}$ & rank ($\omega=0.5$) \\ [0.5ex]
\hline
001 & 1 & 1 &  1 & 1 & 1 & 1 \\
002 & 1 & 1 & 1 & 1 & 0.5 & 0.75 \\
003 & 1 & 1 & 0.5 & 0.75 &0 & 0.375\\
001 & 2 & 1 & 0 & 0.5 &1 & 0.75 \\
001 & 3 & 0 & 0 & 0 &1 & 0.5\\
\hline
\end{tabular}
\label{tb:results}
\end{table}
%\end{example}
